---
title: Scorecard
output: html_document
params: 
  sitename: "5625200827"
  max.activity.date: '2024-08-13'
  max.days: 45
  nightat: "03:00"
  minseglen.global.days: 7
  max.periodic.cpts: 4
  minseglen.periodic: 20
  
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

### credentials
config.folder <- 'connection_strings/'
live <- "config_live.txt"
live <- paste(config.folder, live, sep = '')
#routinecon <- "config_routine.txt"
#routinecon <- paste(config.folder, routinecon, sep = '')
#credentials <- read.csv(paste0(config.folder,'credentials.txt'))

### sources
source('sncirc/R/PGcpt-general.R')
source('sncirc/R/pg_data_gen.R')
source('sncirc/R/SN-circ-general.R')
source('PGCircRoutine.R')

### functions
DialPlots <- 
  # generates x day dial plot 
  function(data,days,tz,nightstart,nightend,daystart,sensor_names) 
  { 

    plot.data <-
    messages |>
  
      dplyr::mutate(ts=as.POSIXct(datecreated,origin='1970-01-01',tz=tz)) |>
      dplyr::mutate(date=as.Date(ts, tz=tz)) |>
      dplyr::filter(dplyr::between(date,max(date)-(days-1),max(date))) 
    
    if(nrow(plot.data)>0){
    
    motion.colours <-
      data.frame(breaks = c('Bathroom','Bedroom','Hallway','Kitchen','Lounge','Toilet','Living Room'),
                 values = c('#88CCEE','#332288','#DDCC77','#117733','#CC6677','#888888','#CC6677'))
      
      
    path.data <-
      plot.data |>
      dplyr::filter(icon == 'MOTN') |>
      dplyr::mutate(ts=as.POSIXct(datecreated,origin='1970-01-01',tz=tz)) |>
      dplyr::mutate(date=as.Date(ts, tz='Europe/London')) |>
      dplyr::mutate(tsgmt=as.POSIXct(howzfunc::dechour(ts)*3600,origin='1970-01-01',tz='GMT')) |>
      dplyr::arrange(ts) |>
      dplyr::mutate(gap = as.integer(ts-dplyr::lag(ts))) |>
      dplyr::mutate(period = ifelse(sensor_name == dplyr::lag(sensor_name) & gap < 600,0,ts)) |>
      dplyr::mutate(period = ifelse(as.integer(difftime(date,dplyr::lag(date)))>0,ts,period)) |>
      na.omit() |>
      dplyr::mutate(period = cummax(period)) |>
      dplyr::mutate(date=as.POSIXct(date)) |>
      dplyr::filter(sensor_name %in% sensor_names) |>
      dplyr::mutate(sensor_name= ifelse(sensor_name== 'Main Door','Main',sensor_name)) |>
      dplyr::mutate(sensor_name= ifelse(sensor_name== 'Back Door','Back',sensor_name)) |>
      dplyr::mutate(sensor_name= ifelse(sensor_name== 'Front Door','Front',sensor_name)) |>
      dplyr::rename(Location=sensor_name) 
        
   
    event.data <-
      plot.data |>
      dplyr::filter(icon != 'MOTN') |>
      dplyr::filter(sensor_name %in% sensor_names) |>   
      dplyr::mutate(bin=as.POSIXct(datecreated %/% 1800 * 1800 + 900)) |>
      dplyr::select(bin,date,sensor_name) |>
      dplyr::group_by(bin) |>
      dplyr::mutate(sensor_name = ifelse(sensor_name == 'Fridge Door','Fridge',sensor_name)) |>
      dplyr::mutate(label = paste(unique(sensor_name), collapse = ' ')) |>
      dplyr::mutate(bingmt=as.POSIXct(howzfunc::dechour(bin)*3600,origin='1970-01-01',tz='GMT')) |>   
      dplyr::group_by(bingmt,date,label) |>
      dplyr::summarise() |>
      dplyr::mutate(date=as.POSIXct(date))

    xlims <- c(as.POSIXct(0,origin='1970-01-01',tz='GMT'),as.POSIXct(86400,origin='1970-01-01',tz=tz))
    ylims <- c(max(plot.data$date)-(days+1),max(plot.data$date)) |> as.POSIXct()
    dates <- data.frame(date=as.POSIXct(as.Date(unique(plot.data$date))),tsgmt=as.POSIXct(0,origin='1970-01-01',tz='GMT')) |>
      dplyr::mutate(label=format(date,'%d/%m')) 

 ggplot2::ggplot()+ 
  ggplot2::coord_polar() +
  ggplot2::geom_text(data=dates,ggplot2::aes(tsgmt,date,label=label),size=8)+
  {if(nrow(path.data)>1) ggplot2::geom_path(data = path.data, ggplot2::aes(tsgmt,date,group=period,color=Location), size=22)}+
  ggplot2::geom_point(data = event.data, ggplot2::aes(bingmt,date), size=6, shape=16, color='black')+
  ggrepel::geom_label_repel(data = event.data, ggplot2::aes(bingmt,date,label=label),size=8) +
  ggplot2::geom_vline(xintercept = nightstart,colour='#117733',linetype='dashed',size=1)+
  ggplot2::geom_vline(xintercept = nightend,colour='#DDCC77',linetype='dashed',size=1)+
  ggplot2::geom_vline(xintercept = daystart,colour='#882255',linetype='dashed',size=1)+
  ggplot2::scale_x_datetime(breaks = scales::date_breaks('1 hour'),labels=scales::label_date('%H:%M'),limits=xlims)+
  ggplot2::scale_y_datetime(breaks = scales::date_breaks('1 day'),limits=ylims)+
  ggplot2::labs(x='',y='')+
  ggplot2::theme_bw()+
  ggplot2::scale_color_manual(breaks = motion.colours$breaks, values = motion.colours$values)+   
  ggplot2::theme(axis.text.y=ggplot2::element_blank(),axis.ticks.y=ggplot2::element_blank(),text = ggplot2::element_text(size = 36,face='bold'),legend.position = 'bottom')

    } else {
      ggplot2::ggplot()+
      ggplot2::labs(title = 'No data')
    }
  }

### define dots
orangefill <-
  '<span style = "color: #DDCC77;;">&#11044;</span>'
greenfill <-
  '<span style = "color: #117733;;">&#11044;</span>'  
greyfill <-
  '<span style = "color: grey;">&#11044</span>' 
orangecircle <-
  '<span style = "color: #DDCC77;">&#9711</span>' 

dots <-
  data.frame(state=c('Within Bounds', 'Out Of Bounds', 'No Sensors','Sensors Offline','No VIP'),
             icon=c(greenfill,orangefill,greyfill,orangecircle,greyfill))

alertnames <-
  data.frame(time=c('Day','Day','Day','Night','Night','24 hrs'),
    concern=c('Hygiene','Nutrition','Social','Rest','Door at night','Circle of care'),
             type=c('Bathroom Day','Food Prep Day','External Door Open Day','Night Activity Ex Bedroom Five Relative','External Door Open Night','Family Using App'))

settings_names <-
  data.frame(key=c('nightstart','nightend','nightlength','daystart','wait_time'),
             parameter=c('Night start','Night end','Length of night','Day start','Wait Time'))

limits <-
    data.frame(key=c('nightstart','nightend','nightlength','daystart','wait_time'),
               mean=c(-0.373,6.9,7.27,9.0,2.35),
               sd=c(1.3,1.1,1,1.5,1),
               settings.distance=c(1.5,1.5,1.5,1.5,1.5))

### get site details 
casequery <- paste0("SELECT name AS Group,group_site_nickname AS Reference,site_nickname AS Circle,Sitename,to_timestamp(min(datecreated))::date AS Since,
to_timestamp(max(datecreated))::date AS Until,
to_timestamp(max(datecreated))::date - to_timestamp(min(datecreated))::date AS days,
now() AT TIME ZONE 'Europe/London' AS Generated FROM
messages 
JOIN howzsites
USING (hid)
LEFT JOIN howzsite_group_mappings
USING (hid)
LEFT JOIN groups
USING (group_id)
LEFT JOIN usersitemappings
USING (hid)
WHERE sitename = ",params$sitename," AND ROLE = 0 GROUP BY name,sitename,group_site_nickname,site_nickname,now()")

case <-
  howzfunc::hsql(live,casequery)

### get alerts
alertsquery <- 
paste0("SELECT DISTINCT sitename,textid,header,body,TO_TIMESTAMP(datecreated)::date AS date,systeminsertdate FROM messages 
JOIN howzsites
USING (hid)
JOIN messagetext
USING (textid)
WHERE textid >= 109
AND datecreated >= (EXTRACT(EPOCH FROM now())-(86400*8))::INTEGER
AND sitename = ", params$sitename)

status <- 
  howzfunc::hsql(live,alertsquery) |>
  tidyr::separate(header,into=c('type','state'),sep='-') |>
  dplyr::mutate(dplyr::across(3:4,stringi::stri_trim_both)) |>
  dplyr::inner_join(dots,by='state') |>
  dplyr::inner_join(alertnames,by='type') |>
  dplyr::group_by(type,date) |> # needed to remove duplicate values created in initial releases
  dplyr::slice_min(order_by = systeminsertdate,n = 1) |>
  dplyr::ungroup()


pquery <- paste0("SELECT sitename,daystart::text,nightstart::text,nightend::text,wait_time::text,systeminsertdate FROM alerts
JOIN howzsites 
USING (hid)
WHERE sitename = ",params$sitename)

settings <-
  howzfunc::hsql(live,pquery) |>
  tidyr::gather(key=key,value=setting,2:5) |>
  dplyr::mutate(setting=as.POSIXct(setting,format='%H:%M:%S')) |>
  dplyr::mutate(setting=howzfunc::dechour(setting)) |>
  dplyr::mutate(setting=ifelse(setting > 12, setting-24, setting)) |>
  tidyr::spread(key=key,value=setting) |>
  dplyr::mutate(nightlength=nightend-nightstart) |>
  tidyr::gather(key=key,value=setting,3:7) |>
  dplyr::inner_join(settings_names,by='key') |>
  dplyr::inner_join(limits,by='key') 

### get activity messages
pquery <- paste0("SELECT h.sitename,datecreated,short_description,sensor_name,macsidid,activity_tracking,icon FROM messages m
  JOIN messagetext USING (textid)
  LEFT JOIN device_howzsite_mappings dhm on m.macsidid = dhm.mac_sid_id
  JOIN howzsites h ON h.hid = m.hid
  LEFT JOIN howzsite_group_mappings hgm ON h.hid = hgm.hid
  LEFT JOIN groups g ON g.group_id = hgm.group_id
  WHERE sitename = ",params$sitename) 

messages <- howzfunc::hsql(live,pquery) |>
  dplyr::filter(activity_tracking) ## |>
  ## dplyr::filter(!(sensor_name %in% c('Bedroom'))) 

max.activity.date = ifelse(is.null(params$max.activity.date),
                            as.Date(as.POSIXct(max(messages$datecreated),origin='1970-01-01',tz='Europe/London')),
                            as.Date(params$max.activity.date))

### get fitted values and dotplots from PGCirc
PGCircOutput <-
  PGCircRoutine(data = messages, 
                tz = 'Europe/London',
                max.days = params$max.days,
                bins = 96,
                max.date = max.activity.date,
                minseglen.periodic = params$minseglen.periodic,
                minseglen.global.days = params$minseglen.global.days,
                max.periodic.cpts = params$max.periodic.cpts,
                pen.global.mult=1,
                wt.confidence = .99,
                plotdays = params$max.days+1,
                nightat = params$nightat,
                title="", #paste0(params$sitename,'-',case$reference),
                subtitle="", #gsub(", ", " ", toString(unique(messages$sensor_name))),
                caption = FALSE
                )

fit <-
  PGCircOutput$summary |>
  dplyr::select(nightstart,nightend,daystart,day.wt) |>
  dplyr::rename(wait_time=day.wt) |>
  tidyr::gather(key=key,value=fit,1:4) |>
  dplyr::mutate(fit=ifelse(fit > 12 & key !='wait_time', fit-24, fit)) 

### put fitted values and settings into a single table and compare differences between fit and average values and fit and settings
settings <-
  settings |>
  dplyr::inner_join(fit,by="key") |>
  dplyr::mutate(fit.mean.distance = abs(mean-fit)/sd) |>
  dplyr::mutate(fit.mean.status=ifelse(fit.mean.distance>2,orangefill,greenfill)) |>
  dplyr::mutate(fit.settings.distance=abs(setting-fit)) |>
  dplyr::mutate(fit.settings.status=ifelse(fit.settings.distance>1.5,orangefill,greenfill)) |>
  dplyr::mutate(fit=as.POSIXct(fit*3600,tz='GMT',origin='1970-01-01')) |>
  dplyr::mutate(fit=format(fit,format='%H:%M')) |>
  dplyr::mutate(setting=as.POSIXct(setting*3600,tz='GMT',origin='1970-01-01')) 

### reformat settings to make it easier to pass to dial plot
plot.settings <-
  settings |>
  dplyr::select(key,setting) |>
  tidyr::spread(key=key,value=setting)

### message re changes in routine

settings.message <- paste(  '<span style = "color: grey;">&#11044;</span>','Settings last updated', format(as.Date(unique(settings$systeminsertdate)),format='%d/%m/%Y'))
days.last.change <- as.integer(Sys.Date()-PGCircOutput$summary$last.change)
change.message <- paste(  '<br> <span style = "color: #DDCC77;">&#11044;</span>','Routine change detected within last 7 days on',
                             format(as.Date(PGCircOutput$summary$last.change),format='%d/%m/%Y'))
no.change.message <- paste(  '<br> <span style = "color: #117733;">&#11044;</span>','Routine stable for',days.last.change,'days since',
                                   format(as.Date(PGCircOutput$summary$last.change),format='%d/%m/%Y'))
routine.message <- ifelse(days.last.change<=7,change.message,no.change.message) 

### placeholder for notes
notes <-
  data.frame(note=c(''))

### logging of results
write.csv(x = PGCircOutput$summary,file = paste0('log/',params$sitename,'-',params$max.periodic.cpts,'.csv'))



```

```{r, echo=FALSE}
### logo
htmltools::a(href='https://www.howz.com',
htmltools::img(src = knitr::image_uri(file.path("/users/jonathanburr/r analysis/howz white.png")), 
               alt = 'logo', 
               style = 'position:absolute; top:0;right:0; padding:10px;')
)
```

### Case

```{r, echo=FALSE}
case |>
  dplyr::mutate(generated=format(generated,format='%d/%m/%Y %H:%M')) |>
  dplyr::mutate(since = format(since,format='%d/%m/%Y')) |>  
  dplyr::mutate(until = format(until,format='%d/%m/%Y')) |>
  #dplyr::select(group,reference,sitename,since,until,days,generated) |>
  kableExtra::kbl(escape=FALSE,align = 'l',col.names=c('Group','Reference','Circle','Sitename','Since','Until','Days','Generated')) |>
  kableExtra::kable_styling(full_width = FALSE,position = 'l') 

```

### Care
Night period statuses are for the night starting at the end of the date shown

```{r, echo=FALSE}
if(nrow(status>1)) {
status |>
  dplyr::select(time,concern,icon,date) |>
  dplyr::rename(Time=time,Concern=concern) |>
  tidyr::spread(key = date, value = icon) |>
  dplyr::rename_with(~ format(as.Date(.x),format='%d/%m'),-1:-2) |>
  kableExtra::kbl(escape=FALSE,align=c('l','l',rep('c',8))) |>
  kableExtra::kable_styling(full_width = TRUE,position = 'l') |>
  kableExtra::collapse_rows(columns = 1,valign = 'top') 
}
```

### Routine
`r settings.message`
`r routine.message`

```{r, echo=FALSE}
settings |>
  dplyr::mutate(setting=format(setting,format='%H:%M')) |>
  dplyr::select(parameter,fit,fit.mean.status,setting,fit.settings.status) |>
  kableExtra::kbl(escape=FALSE,align=c('l','l',rep('c',7)),col.names=c('Parameter','Calculated','Reasonable','Setting','Close to Calculated')) |>
  kableExtra::kable_styling(full_width = FALSE,position = 'l') 

```

---

### Last 7 days by type of activity

#### Motion and external doors

```{r, echo=FALSE}

movement <-
    c('Main Door','Front Door','Back Door',
                      'Bedroom',
                      'Living Room',
                      'Lounge',
                      'Kitchen',
                      'Toilet',
                      'Bathroom',
                      'Hallway',
                      'Landing')

ADL <-
    c('Microwave','Toaster','Grill','Fridge Door','Bathroom','Toilet',
                      'Kettle',
                      'Kitchen')

```



```{r, echo=FALSE, fig.dim = c(24,24)}
DialPlots(data = messages,days = 7,tz = 'Europe/London',nightstart = plot.settings$nightstart,nightend = plot.settings$nightend,daystart = plot.settings$daystart,
          sensor_names = movement)
```

#### Daily Living
Appliance use grouped together in 30 minute periods

```{r, echo=FALSE, fig.dim = c(24,24)}
DialPlots(data = messages,days = 7,tz = 'Europe/London',nightstart = plot.settings$nightstart,nightend = plot.settings$nightend,daystart = plot.settings$daystart,
          sensor_names = ADL)
```

---

### Long term periods of activity and changes 
`r paste0('Date shown up to ',format(as.Date(max.activity.date),format='%d/%m/%Y'),'.  Shades of grey group periods of similar levels of activity and solid vertical lines indicate changes in behaviour.   Dotted horizontal lines are the current settings for night start, night end and day start.  Dotted vertical line is at 6 weeks.')`
```{r, echo=FALSE, fig.dim = c(10, 6)}

PGCircOutput$plot +
  ggplot2::geom_hline(yintercept = plot.settings$nightend,color='orange',linetype='dashed')+
  ggplot2::geom_hline(yintercept = plot.settings$nightstart,color='green',linetype='dashed')+
  ggplot2::geom_hline(yintercept = plot.settings$daystart,color='red',linetype='dashed')
```

#### Routine 
```{r, echo=FALSE, fig.dim = c(10, 6)}
options(knitr.kable.NA = '-')
PGCircOutput$routines |>
  kableExtra::kbl(col.names=c('Date of change','Days','Night start','Night end','Night length','Night start','Night end','Night length','Day','Night'),
                  caption='Values based on routines calculated by the Howz algorithm NOT current settings') |>
  kableExtra::add_header_above(c(" " = 2, "Time" = 2, "Hours"=1, "Change in hours" = 3, "Wait Time" = 2)) |>
  kableExtra::kable_styling(full_width = FALSE,position = 'l') 
```

#### Activity
```{r, echo=FALSE, fig.dim = c(10, 6)}
options(knitr.kable.NA = '-')
PGCircOutput$activity |>
  dplyr::mutate(across(3:6, ~ round(.x * 100,1))) |>
  kableExtra::kbl(col.names=c('Date of change','Days','Day','Night','Day','Night','Ratio'),
                  caption='Values based on routines calculated by the Howz algorithm NOT current settings') |>
  kableExtra::add_header_above(c(" " = 2, "% of time active" = 2, "% change" = 2, " "=1)) |>
  kableExtra::kable_styling(full_width = FALSE,position = 'l') 
```



---
## Current warnings

```{r, echo=FALSE}
status |>
  dplyr::group_by(type) |>
  dplyr::slice_max(order_by = date,n = 1) |>  
  dplyr::filter(state != 'Within Bounds') |>
  dplyr::ungroup() |>
  dplyr::select(time,concern,body) |>
  dplyr::arrange(time,concern) |>
  kableExtra::kbl(escape=FALSE,col.names = c('Time','Concern','Warning')) |>
  kableExtra::kable_styling()
```
