---
title: "Everon Night"
author: "Jonathan Burr"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(ggplot2)
require(dplyr)
options(scipen = 999)
Sys.timezone()
knitr::opts_chunk$set(fig.width = 10, fig.height = 5)

```

## Objective

To assess how the activity data from the Everon pendant can be used for analysis of night periods

## Get data and save locally

```{r}

data <- readRDS('data/everon/everon_processed.rds')

```

## Define night date

Nights are defined by the date before midnight

```{r}
 
data <- 
  data |>
  dplyr::mutate(date = as.Date(as.character(substr(ts, 1,12)))) |>
  dplyr::mutate(night_date = ifelse(lubridate::hour (ts) < 12, date-1, date)) |>
  dplyr::mutate(night_date = as.Date(night_date))                

```


## Truncate for night period 

For this site 12 hours from 9pm to 9am

```{r}
 
data <- 
  data |>
  dplyr::filter((lubridate::hour(ts) >= 22) | (lubridate::hour(ts) < 8))          

```

## Plot values to find suitable nights for analysis
A visual check suggests that night with a mean value > 2 are good candidates for analysis

```{r}

night_mean <-
data |>
  dplyr::group_by(night_date) |>
  dplyr::summarise(mean = mean(value))

data |>
  dplyr::mutate(dechour = howzfunc::dechour(ts)) |>
  dplyr::mutate(dechour = ifelse(dechour < 12, dechour+24, dechour)) |>
  ggplot(aes(dechour,value)) + geom_point() + facet_wrap(night_date ~.) +
  geom_label(data = night_mean, aes(23,30,label = round(mean,2))) +
  labs(title = 'Hours from midday')

```

## Get sleep data from Apple watch
### Compare with categorical data 

```{r}

sleep <- readRDS('sleep.rds') |>
  dplyr::mutate(start_time = lubridate::with_tz(start_time, tzone = 'Europe/London'),
                end_time = lubridate::with_tz(end_time, tzone = 'Europe/London')) |>
  dplyr::mutate(date = as.Date(as.character(substr(start_time, 1,12)))) |>
  dplyr::mutate(night_date = ifelse(lubridate::hour (start_time) > 12, date, as.Date(date-1))) |>
  dplyr::mutate(night_date = as.Date(night_date))

everon_data <-
  night_mean |>
  dplyr::filter(mean > 2) |>
  dplyr::inner_join(data, by = 'night_date')

plot_everon <-
everon_data |>
dplyr::filter(night_date == '2025-05-01') 

plot_sleep <-
  sleep |>
  dplyr::filter(value == 'Awake') |>
  dplyr::filter(night_date == '2025-05-04') 

ggplot() +
  geom_point(data = plot_everon, aes(ts, value)) + facet_grid(night_date ~.) +
  geom_segment(data = plot_sleep, aes(x = start_time, xend = end_time, y= -10, yend = -10, color=value))
  

plot_everon <-
  everon_data |>
  dplyr::filter(night_date == '2025-05-04') |>
  dplyr::filter(value >0) |>
  dplyr::mutate(bin = lubridate::floor_date(ts, unit = '15 minutes')) |>
  dplyr::group_by(bin) |>
  dplyr::summarise(max = mean(value))
  
ggplot() +
  geom_point(data = plot_everon, aes(bin, max)) +
  geom_segment(data = plot_sleep, aes(x = start_time, xend = end_time, y= -10, yend = -10, color=value))  

sum_sleep <-
  sleep |>
  dplyr::mutate(period = as.integer(end_time - start_time)) |>
  dplyr::mutate(period_awake = ifelse(value == 'Awake', period, 0)) |>
  group_by(night_date) |>
  dplyr::summarise(awake = sum(period_awake), start_night = min(start_time), end_night = max(end_time))

sum_everon <-
  everon_data |>
  dplyr::inner_join(sum_sleep, by = 'night_date') |>
  dplyr::filter(ts >= start_night & ts <= end_night) |>
  dplyr::filter(value >0) |>
  dplyr::mutate(bin = lubridate::floor_date(ts, unit = '15 minutes')) |>
  dplyr::group_by(night_date, bin) |>
  dplyr::summarise(mean = mean(value)) |>
  dplyr::filter(mean >= 3) |>
  dplyr::group_by(night_date) |>
  dplyr::summarise(n = n())
  
comp <-  
sum_everon |>
  dplyr::inner_join(sum_sleep, by ='night_date') #|>
  #filter(n<20) 

model <- lm(n ~ awake, comp)
corr <- summary(model)$r.squared

comp |>
ggplot(aes(n, awake)) + geom_point() + geom_smooth(method = 'lm') +
labs(title = 'Comparison of Everon and Apple Watch data', subtitle = paste('R^2 =',round(corr,3)), x = 'Everon - Number of above threshold bins', y = 'Apple Watch seconds awake')  

```

### Compare with energy data

```{r}

active_energy <- readRDS('active_energy.rds') |>
  dplyr::mutate(start_time = lubridate::with_tz(start_time, tzone = 'Europe/London'),
                end_time = lubridate::with_tz(end_time, tzone = 'Europe/London')) |>
  dplyr::mutate(date = as.Date(as.character(substr(start_time, 1,12)))) |>
  dplyr::mutate(night_date = ifelse(lubridate::hour (start_time) > 12, date, as.Date(date-1))) |>
  dplyr::mutate(night_date = as.Date(night_date))

active_energy <-
  active_energy |>
    dplyr::mutate(bin = lubridate::floor_date(start_time, unit = '15 minutes')) 

bin_active_energy <-
  active_energy |>
  dplyr::group_by(bin) |>
  dplyr::summarise(active_energy=sum(value))

sum_active_energy <-
  active_energy |>
  dplyr::inner_join(sum_sleep, by = 'night_date') |>
  dplyr::filter(start_time >= start_night & end_time <= end_night) |>
  dplyr::group_by(night_date) |>
  dplyr::summarise(active_energy = sum(value)) 

comp <-  
sum_everon |>
  dplyr::inner_join(sum_active_energy, by ='night_date') #|>
  #filter(n<20) 

model <- lm(n ~ active_energy, comp)
corr <- summary(model)$r.squared

comp |>
ggplot(aes(n, active_energy)) + geom_point() + geom_smooth(method = 'lm') +
labs(title = 'Comparison of Everon and Apple Watch data', subtitle = paste('R^2 =',round(corr,3)), x = 'Everon - Number of above threshold bins', y = 'Apple Watch seconds awake')  


comp <-  
everon_data |>
  dplyr::filter(value >0) |>
  dplyr::mutate(bin = lubridate::floor_date(ts, unit = '15 minutes')) |>
  dplyr::group_by(bin) |>
  dplyr::summarise(mean = mean(value)) |>
  dplyr::filter(mean > 3) |>
  dplyr::inner_join(bin_active_energy, by ='bin')

model <- lm(mean ~ active_energy, comp)
corr <- summary(model)$r.squared
readr::write_csv(comp, 'comp.csv')


comp |>
ggplot(aes(mean, active_energy)) + geom_point() + geom_smooth(method = 'lm') +
labs(title = 'Comparison of Everon and Apple Watch data', subtitle = paste('R^2 =',round(corr,3)), x = 'Everon - Number of above threshold bins', y = 'Apple Watch seconds awake') 

```

## Calculation of rest start and end times
### Find changepoints

```{r}

cpts <- 
  data$value |>
  changepoint::cpt.mean(method = 'PELT', minseglen = 480)

data$cpt[attr(cpts,'cpts')] <- TRUE

cpt_data <-
data |>
  dplyr::mutate(cpt = row_number(cpt)) |>
  dplyr::mutate(cpt = zoo::na.locf(cpt, na.rm= FALSE)) |>
  dplyr::mutate(cpt = replace(cpt, is.na(cpt),0)) |>
  dplyr::group_by(cpt) |>
  #dplyr::mutate(mean = ifelse(value == 0, 0, mean(value)), var = ifelse(value == 0, 0, var(value)))
  dplyr::mutate(mean = mean(value), var = var(value))
  
cpt_data |>
  dplyr::filter(night_date >= '2025-05-02' & night_date <= '2025-05-03') |>
  ggplot() +
  geom_point(aes(ts, value)) +
  geom_step(aes(ts,mean))             
              
cpt_data_segments <-
  cpt_data |>
  dplyr::group_by(mean) |>
  dplyr::summarise()

kmeans(cpt_data_segments$mean,centers = 2)

```




