---
title: "Everon Investigation - 1"
author: "Jonathan Burr"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(ggplot2)
require(dplyr)

# Source function 
source("~/LocalRepo/Everon/PGCIRC/PGcpt.R")
source("~/LocalRepo/Everon/PGCIRC/SNcirc.R")
options(scipen = 999)


```

## Objective

To assess whether the activity data from the Everon pendant can be used
with the PGCirc algorithm

## Get data and save locally

Everon upload data for the devices daily at 9am at the url below.

```{r}

url <- "https://howzdata.blob.core.windows.net/upload/influx_export.csv"
destfile <- "/Users/jonathanburr/LocalRepo/Everon/data/everon/influx_export.csv"
download.file(url, destfile, method = "curl")
Sys.timezone()
knitr::opts_chunk$set(fig.width = 10, fig.height = 5)

```

## Define parameters

```{r}
 
bins <- 96
bin.size <- "15 minutes"
start.date <- as.POSIXct('2025-04-23 00:00:00')
```

## Get and filter data for specific device

```{r}
data <- readr::read_csv("/Users/jonathanburr/LocalRepo/Everon/data/everon/influx_export.csv") |>
  dplyr::mutate(ts = as.POSIXct(time/1000), tz ='Europe/London') |>
  dplyr::rename(device = tags) |>
  dplyr::mutate(device = gsub(pattern = 'deviceId=','',device)) |>
  dplyr::filter(device == '05301BE3') |>
  dplyr::filter(ts >= start.date)
tail(data)  
nrow(data)
```

## Plot data

```{r}
data |>
  ggplot(aes(ts,value)) + geom_point() 

data |>
  dplyr::summarise(first_obs = min(ts), last_obs = max(ts))
```

## Calculate Out of Range periods

### Overall out of range

```{r}

out.of.range <-
data |>
  dplyr::mutate(gap = as.integer(ts - dplyr::lag(ts))) |>
  dplyr::filter(gap > 1) |>
  dplyr::mutate(date = as.Date(ts)) |>
  dplyr::group_by(date) |>
  dplyr::summarise(n = n(), mean = mean(gap)) 

out.of.range |>
  ggplot(aes(date,n/1440)) + geom_point()
```

### Out of range when physically close

```{r}
out.of.range |>
  filter(date == '2025-05-17')
```

## Create empty set of 1 minute values

When the device is out of range no data is sent which are the blankc
values in the plot above. A full set of timestamps for the observation
period is created to readily determine periods when the device is out of
range.

```{r}

start <-  start.date
end <- max(data$ts)

full_time <- data.frame(ts = seq(start, end-60, by = "1 min")) 

data <-
  data |>
  dplyr::select(ts, value) |>
  dplyr::right_join(full_time, by = 'ts') |>
  dplyr::mutate(value = tidyr::replace_na(value, 0)) |>
  dplyr::arrange(ts)

```

## Clustering of values

```{r}

clustered <-
data |>
  dplyr::filter(ts >= start.date) 

clustered.non.zero <-  ### removes zero (out of range) values from cpt
  clustered |>
  dplyr::filter(value > 0)

cpts <- 
  clustered.non.zero$value |>
  changepoint::cpt.mean(method = 'PELT', minseglen = 5)

clustered.non.zero$cpt[attr(cpts,'cpts')] <- TRUE

clustered <-
  clustered |>
  dplyr::left_join(dplyr::select(clustered.non.zero, -value), by = 'ts') |>
  dplyr::mutate(cpt = row_number(cpt)) |>
  dplyr::mutate(cpt = zoo::na.locf(cpt, na.rm= FALSE)) |>
  dplyr::mutate(cpt = replace(cpt, is.na(cpt),0)) |>
  dplyr::group_by(cpt) |>
  dplyr::mutate(mean = ifelse(value == 0, 0, mean(value)), var = ifelse(value == 0, 0, var(value))) |>
  dplyr::filter(ts >= as.POSIXct('2025-05-18 00:00:00') & ts <= as.POSIXct('2025-05-20 00:00:00')) |>
  dplyr::ungroup()

clustered |>
  ggplot(aes(ts, value)) + geom_point() +
  geom_step(aes(ts, mean), color='red') +
  scale_x_datetime(
    breaks = seq(from = min(clustered$ts), to = max(clustered$ts), by = "240 min"),
    date_labels = "%H:%M",
    timezone = 'Europe/London'
  ) +
  theme(legend.position = "none") 

```

## Group data into 15 minute bins

```{r}
data <-
  data |>
  dplyr::arrange(ts) |>
  dplyr::mutate(bin = lubridate::floor_date(ts, "15 minutes")) |>  
  dplyr::mutate(bin = hms::as_hms(bin)) |>
  dplyr::mutate(date = as.Date(ts)) 
nrow(data)
tail(data)
readr::write_csv(data,'test.csv')
```

## Calculate the average activity value for each bin and day

```{r}
bybin <-
  data |>
  dplyr::group_by(date, bin) |>
  dplyr::summarise(avgbin = mean(value)) |>
  dplyr::ungroup()
readr::write_csv(bybin,'bin.csv')

```

## Run PGCirc

```{r}
dat = data_circ(bybin$avgbin,96)
results = pgcpt(dat, period.len=bins, minseglen.periodic=4, minseglen.global= 7*96,
                method.periodic="SNcirc", method.global="PELT", max.periodic.cpts=10, penalty.periodic=3*log(length(dat)),
                penalty.global=1*log(nrow(dat)), dist="Normal mean", restrict=TRUE, circData=FALSE)

length(dat)
tail(dat)
results$pgcpt.results$Periodic_cpt[[1]]

cpts <- data.frame(cpts = as.POSIXct((results$pgcpt.results$Periodic_cpt[[1]]-1) * 900))
cpts
```

### Add lines from PGCirc to plot

```{r}
class(bybin$bin)

bybin |>
  dplyr::group_by(bin) |>
  dplyr::summarise(avgbin = mean(avgbin)) |>
  ggplot2::ggplot(aes(bin, avgbin)) + geom_bar(stat='identity') +
  ggplot2::geom_vline(xintercept = cpts$cpts)
```

### Create synthetic data to demonstrate change detection

### Comparison of sleep data

```{r}

sleep <- readRDS('sleep.rds')

sleep_start <- as.POSIXct('2025-05-13 22:00:00')
sleep_end <- as.POSIXct('2025-05-14 7:30:00')

sleep_episodes <-
  sleep |>
  dplyr::filter(start_time >= sleep_start & start_time <= sleep_end) |>
  dplyr::mutate(value = as.factor(value)) |>
  dplyr::mutate(value = forcats::fct_relevel(value, "AsleepDeep", "AsleepCore", "AsleepREM", "Awake", "Energy", "Everon"))

everon_data <-
  data |>
  dplyr::filter(ts >= sleep_start & ts <= sleep_end) |>
  dplyr::arrange(ts) |>
  dplyr::mutate(value = (value-min(value))/(max(value)-min(value))) |>
  howzfunc::howzcp(seglen = 5, column = 2)

ggplot() +
  geom_segment(data = sleep_episodes, aes(x = start_time, xend = end_time, y = value, yend = value), linewidth = 5) +
  geom_segment(data = energy_episodes, aes(x = start_time, xend = end_time, y = 'Energy', yend = 'Energy', color = scaled_value, linewidth = scaled_value)) +
  geom_line(data = everon_data, aes(x = ts, y = 'Everon', linewidth = mean)) +
  scale_x_datetime(
    breaks = seq(from = min(everon_data$ts), to = max(everon_data$ts), by = "60 min"),
    date_labels = "%H:%M",
    timezone = 'Europe/London'
  ) +
  theme(legend.position = "none") 

```

### Comparison of energy data - need to try binning and compare bins

```{r}

energy <- readRDS('active_energy.rds') |>
  dplyr::mutate(duration = as.numeric(end_time - start_time)) |>
  dplyr::mutate(energy_rate = as.numeric(value) / duration)

energy_episodes <- 
  energy |> dplyr::filter(start_time >= as.POSIXct('2025-05-13 22:00:00') & start_time <= as.POSIXct('2025-05-14 07:30:00')) |>
  dplyr::mutate(value = as.numeric(value)) |>
  dplyr::mutate(scaled_value = ((energy_rate - min(energy_rate))) / (max(energy_rate)- min(energy_rate)))

clustered_scaled <-
  clustered |>
  dplyr::filter(ts >= as.POSIXct('2025-05-18 13:00:00') & ts <= as.POSIXct('2025-05-24 07:30:00')) |>
    dplyr::mutate(scaled_value = ((mean - min(mean))) / (max(mean)- min(mean))) 

ggplot() + 
geom_segment(data = energy_episodes, aes(x = start_time, xend = end_time, y=scaled_value, yend=scaled_value), linewidth = 5, color = 'red') +  
geom_point(data = clustered_scaled, aes(x = ts, y = scaled_value)) +
theme(legend.position = "none")


```
