---
title: "Everon Routine"
author: "Jonathan Burr"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(ggplot2)
require(dplyr)

# Source function 
source("~/LocalRepo/Everon/PGCIRC/PGcpt.R")
source("~/LocalRepo/Everon/PGCIRC/SNcirc.R")
options(scipen = 999)


```

## Objective

To assess whether the activity data from the Everon pendant can be used
with the PGCirc algorithm

## Get data and save locally

Everon upload data for the devices daily at 9am at the url below.

```{r}

url <- "https://howzdata.blob.core.windows.net/upload/influx_export.csv"
destfile <- "/Users/jonathanburr/LocalRepo/Everon/data/everon/influx_export.csv"
download.file(url, destfile, method = "curl")
Sys.timezone()
knitr::opts_chunk$set(fig.width = 10, fig.height = 5)

```

## Define parameters

```{r}
 
bins <- 96
bin.size <- "15 minutes"
start.date <- as.POSIXct('2025-04-23 00:00:00')
```

## Get and filter data for specific device

```{r}
data <- readr::read_csv("/Users/jonathanburr/LocalRepo/Everon/data/everon/influx_export.csv") |>
  dplyr::mutate(ts = as.POSIXct(time/1000), tz ='Europe/London') |>
  dplyr::rename(device = tags) |>
  dplyr::mutate(device = gsub(pattern = 'deviceId=','',device)) |>
  dplyr::filter(device == '05301BE3') |>
  dplyr::filter(ts >= start.date)
tail(data)  
nrow(data)
```

## Plot data

```{r}
data |>
  ggplot(aes(ts,value)) + geom_point() 

data |>
  dplyr::summarise(first_obs = min(ts), last_obs = max(ts))
```

## Calculate Out of Range periods

### Overall out of range

```{r}

out.of.range <-
data |>
  dplyr::mutate(gap = as.integer(ts - dplyr::lag(ts))) |>
  dplyr::filter(gap > 1) |>
  dplyr::mutate(date = as.Date(ts)) |>
  dplyr::group_by(date) |>
  dplyr::summarise(n = n(), mean = mean(gap)) 

out.of.range |>
  ggplot(aes(date,n/1440)) + geom_point()
```

### Out of range when physically close

```{r}
out.of.range |>
  filter(date == '2025-05-17')
```

## Create empty set of 1 minute values

When the device is out of range no data is sent which are the blankc
values in the plot above. A full set of timestamps for the observation
period is created to readily determine periods when the device is out of
range.

```{r}

start <-  start.date
end <- max(data$ts)

full_time <- data.frame(ts = seq(start, end-60, by = "1 min")) 

data <-
  data |>
  dplyr::select(ts, value) |>
  dplyr::right_join(full_time, by = 'ts') |>
  dplyr::mutate(value = tidyr::replace_na(value, 0)) |>
  dplyr::arrange(ts)

```

## Group data into 15 minute bins

```{r}
data <-
  data |>
  dplyr::arrange(ts) |>
  dplyr::mutate(bin = lubridate::floor_date(ts, "15 minutes")) |>  
  dplyr::mutate(bin = hms::as_hms(bin)) |>
  dplyr::mutate(date = as.Date(ts)) 
nrow(data)
tail(data)
saveRDS(data,'data/everon/everon_processed.rds')
```

## Calculate the average activity value for each bin and day

```{r}
bybin <-
  data |>
  dplyr::filter(date >= '2025-05-01' & date <= '2025-05-07') |>
  dplyr::group_by(date, bin) |>
  dplyr::summarise(avgbin = mean(value)) |>
  dplyr::ungroup()
readr::write_csv(bybin,'bin.csv')

```

## Run PGCirc

```{r}
dat = data_circ(bybin$avgbin,96)
results = pgcpt(dat, period.len=bins, minseglen.periodic=4, minseglen.global= 96,
                method.periodic="SNcirc", method.global="PELT", max.periodic.cpts=10, penalty.periodic=3*log(length(dat)),
                penalty.global=1*log(nrow(dat)), dist="Normal mean", restrict=TRUE, circData=FALSE)

length(dat)
tail(dat)
results$pgcpt.results$Periodic_cpt[[1]]

cpts <- data.frame(cpts = as.POSIXct((results$pgcpt.results$Periodic_cpt[[1]]-1) * 900))
cpts
```

### Add lines from PGCirc to plot

```{r}
class(bybin$bin)

bybin |>
  dplyr::group_by(bin) |>
  dplyr::summarise(avgbin = mean(avgbin)) |>
  ggplot2::ggplot(aes(bin, avgbin)) + geom_bar(stat='identity') +
  ggplot2::geom_vline(xintercept = cpts$cpts)
```

### Dotplot

```{r}

dot_data <-
  data |>
  dplyr::filter(date >= '2025-05-01' & date <= '2025-05-07') 

ggplot() +
  geom_segment(
    data = dot_data,
    aes(
      x = date,
      xend = date,
      y = bin,
      yend = bin + lubridate::duration(15, units = "minutes"),
      linewidth = value
    )
  )

```