---
title: "Sleep Analysis"
author: "Jonathan Burr"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
require(readr)
require(ggplot2)
knitr::opts_chunk$set(fig.width = 10, fig.height = 5)

threshold <- 1
roll_size <- 5
bin_size <- '5 minutes'
msl <- 15


is_night_time <- function(timestamp, night_start, night_end) {
  time_only <- hms::as_hms(timestamp)

  night_start <- hms::as_hms(night_start)
  night_end <- hms::as_hms(night_end)

  if (night_end < night_start) {
    # Overnight case: e.g., 23:00 to 06:00
    return(time_only >= night_start || time_only <= night_end)
  } else {
    # Same-day case: e.g., 20:00 to 23:00
    return(time_only >= night_start && time_only <= night_end)
  }
}

relative_dechour <- function(dechour) {
  ifelse(dechour > 12, dechour - 24, dechour)
}

```


## Objective
Quantity of rest
Quality of rest

## Read in Everon data 

```{r}
everon <-readr::read_csv2("/Users/jonathanburr/LocalRepo/Everon/data/everon/to_postgresql.csv") |>
  dplyr::filter(device_id == 18258) |>
  dplyr::rename(ev_device_id = device_id) |>
  dplyr::mutate(ts = as.POSIXct(datecreated, tz = 'Europe/London')) 

full_time <- data.frame(ts = seq(min(everon$ts), max(everon$ts)-60, by = "1 min")) 

everon <-
  everon |>
  dplyr::right_join(full_time, by = 'ts') |>
  dplyr::mutate(filled = is.na(value)) |>  
  dplyr::arrange(ts) |>
  dplyr::mutate(value = ifelse(is.na(value), 1, value)) |>
  dplyr::mutate(hid = zoo::na.locf(hid), ev_device_id = zoo::na.locf(ev_device_id)) |>
  dplyr::select(-datecreated) |>
  dplyr::mutate(value = log(value+1))

everon <-
  everon |>
  dplyr::mutate(date = as.Date(ts)) |>
  dplyr::mutate(night_of = as.Date(ts - 43200)) |>
  dplyr::mutate(bin = lubridate::floor_date(ts, unit = bin_size)) |>
  dplyr::mutate(rollmean = zoo::rollmean(x= value, k = roll_size, fill = NA, align = 'right')) |>
  dplyr::mutate(rollmedian = zoo::rollmedian(x= value, k = roll_size, fill = NA, align = 'right')) |>
  dplyr::mutate(rollsum = zoo::rollsum(x= value, k = roll_size, fill = NA, align = 'right')) |>
  dplyr::mutate(dechour = relative_dechour(howzfunc::dechour(ts)))

cpts <- changepoint::cpt.mean(data = na.omit(everon$value), method = 'PELT', minseglen = msl)

everon$cpt[attr(cpts,'cpts')] <- TRUE

everon <-
everon |>
  dplyr::mutate(cpt = dplyr::row_number(cpt)) |>
  dplyr::mutate(cpt = zoo::na.locf(cpt, na.rm= FALSE)) |>
  dplyr::mutate(cpt = replace(cpt, is.na(cpt),0)) |>
  dplyr::group_by(cpt) |>
  dplyr::mutate(cptmean = mean(value), cptvar = var(value))





```

## Read in watch data

```{r}

watch <- readr::read_csv("/Users/jonathanburr/LocalRepo/Everon/data/apple_health_export/sleep_summaries.csv",
  col_types = readr::cols(start_time = col_time(format = "%H:%M:%S"),
  end_time = col_time(format = "%H:%M:%S"),
  time_asleep = col_time(format = "%H:%M:%S"),
  awake = col_time(format = "%H:%M:%S"),
  REM = col_time(format = "%H:%M:%S"),
  core = col_time(format = "%H:%M:%S"),
  deep = col_time(format = "%H:%M:%S"),
  sleep_score = col_integer())) |>
  dplyr::rename(wt_device_id = device_id) |>
  dplyr::filter(night_of != '2025-04-29')

```

### Measured nights

```{r}

measured_nights <- unique(watch$night_of)

```


### Correct 
Problem that when device is not being worn will give values of 1 
This will mess with finding night time values
Need to assume that person is active when not being worn
Uses kmeans to find not being worn segment and then changes the data 

```{r}

selected_nights <-
everon |>
  dplyr::ungroup() |>
  dplyr::filter(night_of %in% measured_nights) |>
  na.omit()

k_model <- kmeans(x = dplyr::select(selected_nights, cptmean, cptvar), centers = 3)
selected_nights$cluster <- k_model$cluster

selected_nights |>
  ggplot(aes(dechour,value)) + 
  geom_point() +
  geom_step(aes(dechour, cptmean), color='red') +
  facet_wrap(night_of~.)
  
segments <-
  selected_nights |>
  dplyr::group_by(night_of, cpt) |>
  dplyr::summarise(start = min(dechour), end = max(dechour), cluster = mean(cluster))

selected_nights |>
  ggplot() +
  #geom_point(aes(dechour, as.character(night_of), size = value)) +
  geom_segment(data=segments, aes(x=start, xend = end, y=as.character(night_of), yend=as.character(night_of),color=as.factor(cluster)), size = 5) +
  xlim(-3,9)
  


```





## Join together 

```{r}
combined <-
  watch |>
  dplyr::select(hid, wt_device_id, night_of, start_time, end_time) |>
  dplyr::inner_join(everon, by = c('hid','night_of')) 

combined <-
  combined |>
  dplyr::rowwise() |>
  dplyr::mutate(
    is_night = is_night_time(ts, start_time, end_time)
  ) |>
  dplyr::ungroup() 
                
```  

## Completeness

```{r}

completeness <-
combined |>
  dplyr::filter(is_night) |>
  dplyr::group_by(night_of, wt_device_id) |>
  dplyr::summarise(start = min(ts), end = max(ts), duration = as.numeric(max(ts) - min(ts))*60, readings = dplyr::n(), coverage = readings/duration) 

head(completeness)
```

## Summarise
### Using device boundaries

``` {r}

thresholded <-
  combined |>
  dplyr::filter(is_night) |>
  dplyr::group_by(hid, wt_device_id, night_of, bin) |>
  dplyr::summarise(value = mean(value)) |>
  dplyr::mutate(oob = as.integer(value > threshold)) 

thresholded_sum <-
  thresholded |>
  #dplyr::filter(night_of != '2025-05-27') |>
  #dplyr::filter(night_of != '2025-05-23') |>
  #dplyr::filter(night_of != '2025-05-22') |>
  #dplyr::filter(night_of != '2025-05-14') |>
  dplyr::group_by(hid, night_of, wt_device_id) |>
  dplyr::summarise(oob = sum(oob)) |>
  dplyr::ungroup() |>
  dplyr::inner_join(watch, by = c('night_of', 'hid', 'wt_device_id')) 

# Fit linear model 
apple_model <- lm(howzfunc::dechour(awake) ~ oob, data = dplyr::filter(thresholded_sum, wt_device_id == 18262))
fitbit_model <- lm(howzfunc::dechour(awake) ~ oob, data = dplyr::filter(thresholded_sum, wt_device_id == 18264))

# Get R-squared

thresholded_sum |>
ggplot(aes(oob, awake, color=as.character(wt_device_id))) + geom_point()  + geom_smooth(method = 'lm',se = FALSE) +
geom_label(aes(oob, awake, label = night_of),size =3) + facet_wrap(wt_device_id~., scales = 'free_y') +
  labs(subtitle = paste('Apple:', summary(apple_model)$r.squared, 'Fitbit:',summary(fitbit_model)$r.squared))

```

## Devices against each other

```{r}

apple_only <-
watch |>
  dplyr::filter(wt_device_id == 18262)

fitbit_only <-
watch |>
  dplyr::filter(wt_device_id == 18264)

both_devices <-
  apple_only |>
  dplyr::inner_join(fitbit_only, by = 'night_of') |>
  dplyr::mutate(dplyr::across(dplyr::where(hms::is.hms), howzfunc::dechour)) |>
  dplyr::mutate(dplyr::across(dplyr::matches("start_time|end_time"), relative_dechour)) |>
  dplyr::mutate(length.x = end_time.x - start_time.x) |>
  dplyr::mutate(length.y = end_time.y - start_time.y) |>
  dplyr::mutate(awake_prop.x = awake.x/length.x) |>
  dplyr::mutate(awake_prop.y = awake.y/length.y) #|>
  #dplyr::filter(night_of !=  '2025-05-27')

start_model <- lm(start_time.x ~ start_time.y, data = both_devices)
both_devices |>
  ggplot(aes(start_time.x, start_time.y)) + 
  geom_point() + 
  geom_smooth(method = 'lm', se = FALSE) +
  labs(title = 'start',subtitle = paste('R^2:', summary(start_model)$r.squared)) 

end_model <- lm(end_time.x ~ end_time.y, data = both_devices)
both_devices |>
  ggplot(aes(end_time.x, end_time.y)) +
  geom_point() + 
  geom_smooth(method = 'lm', se = FALSE) +
  labs(title = 'end', subtitle = paste('R^2:', summary(end_model)$r.squared)) +
  geom_label(aes(end_time.x, end_time.y, label = night_of),size =3)

length_model <- lm(length.x ~ length.y, data = both_devices)
both_devices |>
  ggplot(aes(length.x, length.y)) + 
  geom_point() + 
  geom_smooth(method = 'lm', se = FALSE) +
  labs(title = 'length', subtitle = paste('R^2:', summary(length_model)$r.squared)) + geom_label(aes(length.x, length.y, label = night_of),size =3)

awake_model <- lm(awake.x ~ awake.y, data = both_devices)
both_devices |>
  ggplot(aes(awake.x, awake.y)) + 
  geom_point() + 
  geom_smooth(method = 'lm', se = FALSE) +
  labs(title = 'awake',subtitle = paste('R^2:', summary(awake_model)$r.squared)) + geom_label(aes(awake.x, awake.y, label = night_of),size =3)

awake_prop_model <- lm(awake_prop.x ~ awake_prop.y, data = both_devices)
both_devices |>
  ggplot(aes(awake_prop.x, awake_prop.y)) + 
  geom_point() + 
  geom_smooth(method = 'lm', se = FALSE) +
  labs(title = 'awake_prop',subtitle = paste('R^2:', summary(awake_prop_model)$r.squared)) + geom_label(aes(awake_prop.x, awake_prop.y, label = night_of),size =3)


```


```






## Fitbit Investigation

```{r}

test <-
combined |>
  dplyr::filter(is_night) |>
  dplyr::filter(bin >= '2025-05-27 21:00:00' & bin <= '2025-05-28 09:00:00') 
  
test |>  
  ggplot(aes(ts, value, color = value > threshold)) + geom_point() +
  geom_line(aes(ts,rollsum),color='red') +
  labs(title = paste('rollsum', max(test$night_of)), subtitle = paste(format(min(test$ts),'%H:%M'), format(max(test$ts),'%H:%M')))

test |>  
  ggplot(aes(ts, value, color = value > threshold)) + geom_point() +
  geom_line(aes(ts,rollmedian),color='red') +
  labs(title = paste('rollmedian', max(test$night_of)), subtitle = paste(format(min(test$ts),'%H:%M'), format(max(test$ts),'%H:%M')))
  
test |>  
  ggplot(aes(ts, value, color = value > threshold)) + geom_point() +
  geom_line(aes(ts,cptmean),color='red') +
  labs(title = paste('cptmean', max(test$night_of)), subtitle = paste(format(min(test$ts),'%H:%M'), format(max(test$ts),'%H:%M')))              

```

