---
title: "Everon Investigation"
author: "Jonathan Burr"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(ggplot2)

# Source function 
source("~/LocalRepo/Everon/PGCIRC/PGcpt.R")
source("~/LocalRepo/Everon/PGCIRC/SNcirc.R")

```

## Objective 

To assess whether the activity data from the Everon pendant can be used with the PGCirc algorithm

## Get data and save locally

Everon upload data for the devices daily at 4am at the url below.

```{r}

url <- "https://howzdata.blob.core.windows.net/upload/influx_export.csv"
destfile <- "/Users/jonathanburr/LocalRepo/Everon/data/everon/influx_export.csv"
download.file(url, destfile, method = "curl")

```

## Define parameters

```{r}
start <-  as.POSIXct('2025-04-23 00:00:00')
end <- as.POSIXct('2025-05-02 04:00:00')   
bins <- 96
bin.size <- "15 minutes"
```

## Get and filter data for specific device 

```{r}
data <- readr::read_csv("/Users/jonathanburr/LocalRepo/Everon/data/everon/influx_export.csv") |>
  dplyr::mutate(ts = as.POSIXct(time/1000)) |>
  dplyr::rename(device = tags) |>
  dplyr::mutate(device = gsub(pattern = 'deviceId=','',device)) |>
  dplyr::filter(device == '05301BE3')
  
head(data)  
```

## Plot data 

```{r}
data |>
  ggplot(aes(ts,value)) + geom_point() 
```


## Create empty set of 1 minute values

When the device is out of range no data is sent which are the blankc values in the plot above.   A full set of timestamps for the observation period is created to readily determine periods when the device is out of range.

```{r}
full_time <- data.frame(ts = seq(start, end, by = "1 min")) 
```


## Group data into 15 minute bins

```{r}
data <-
  data |>
  dplyr::select(ts, value) |>
  dplyr::right_join(full_time, by = 'ts') |>
  dplyr::mutate(value = tidyr::replace_na(value, 0)) |>
  dplyr::arrange(ts) |>
  dplyr::mutate(bin = lubridate::floor_date(ts, "15 minutes")) |>   
  dplyr::mutate(date = as.Date(bin)) |>
  dplyr::group_by(date) |>
  dplyr::mutate(binnum = dplyr::dense_rank(bin)) |>
  dplyr::ungroup()

tail(data)
```

## Calculate the average activity value for each bin and day

```{r}
bybin <-
  data |>
  dplyr::group_by(date, binnum) |>
  dplyr::summarise(avgbin = mean(value)) |>
  dplyr::ungroup()
```

## Run PGCirc

```{r}
dat = data_circ(bybin$avgbin,96)
results = pgcpt(dat, period.len=bins, minseglen.periodic=8, minseglen.global= 7*96,
                method.periodic="SNcirc", method.global="PELT", max.periodic.cpts=10, penalty.periodic=3*log(length(dat)),
                penalty.global=1*log(nrow(dat)), dist="Normal mean", restrict=TRUE, circData=FALSE)

length(dat)
tail(dat)
results$pgcpt.results$Periodic_cpt[[1]]

```

### Add lines from PGCirc to plot

```{r}
bybin |>
  dplyr::group_by(binnum) |>
  dplyr::summarise(avgbin = mean(avgbin)) |>
  ggplot2::ggplot(aes(binnum, avgbin)) + geom_bar(stat='identity') +
  ggplot2::geom_vline(xintercept = as.integer(results$pgcpt.results$Periodic_cpt[[1]]))
```
